### 2019-07-17 manuscript/abstract.tex
bmc style, background and disproofs written

\documentclass{article}
\usepackage{standalone}

\begin{document}

Active learning for systematic review screening promises to reduce the human effort required to identify relevant documents for a systematic review. Machines and humans  work together, with the human providing training data, and the machine optimising the documents a human screens. However, current approaches lack robust stopping criteria, meaning that such systems are hard to implement in live reviews. This paper introduces a workflow for working with robust and flexible statistical stopping criteria, that offer real work reductions on the basis of a given confidence level of reaching a given recall.

\end{document}